Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-12-08 Manuscript/manuscript.tex
New isocontour figure with a slighlty improved caption. Not final, will update it with more smooth contours.

\begin{abstract}
  % insert abstract here
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. Current estimates of binding affinities and
stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
polarization as a mean field correction, known as the electronic continuum correction (ECC).
This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
in agreement with the previous interpretations of the experimental spectroscopic data.
The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities, suggesting more complex binding stoichiometries than those from simple models
used to interpret the NMR data previously.
The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
of complex biochemical systems at cellular membranes.
\end{abstract}

@@

\section{Introduction}
Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}. Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}. For example, Ca$^{2+}$  is crucial in propagation of neural signals. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}. Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.

Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}. The details of ion binding are, however, not fully consistent in the literature. Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers. In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}. Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}. Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.

Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements. A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}. Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results. 
Analyzing massive amounts of data collected within an open collaboration approach, 
it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.

In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}. Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model. This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}. 
The newly developed ECC-POPC model reproduces the experimentally
measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.

@@

\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations. In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}. Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}. However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.

According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}. Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
\begin{equation}
Q^{ECC} = f_q \cdot Q
\end{equation}
for the ECC corrected charges. 
Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$. This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
It is important to note that the value of the
high frequency dielectric constant 
is around 2 for almost any biologically relevant environment \cite{leontyev11}.
The dielectric discontinuity in a lipid bilayer thus arises only
from the nuclear (orientational) polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
Therefore, the same correction for the electronic polarizability can be 
applied throughout the lipid bilayer/aqueous solution interface.

While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case. Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables. Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}. Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}. Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$. Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).

Here we develop a ECC-POPC lipid model that accurately describes binding 
of sodium and calcium ions to the POPC  lipid bilayer. 
The Lipid14~\cite{dickson14} force field 
(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
point since 
it provides the most accurate response of the head groups to ions among the available 
lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
provides realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
We applied the ECC correction 
to the Lipid14 model of POPC by scaling 
partial charges of the head group, glycerol 
backbone, and carbonyl regions. 
These are the polar parts of phospholipids which can 
contribute to the cation binding.

To reproduce the experimental ion binding affinities,
we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
The ion binding affinity was benchmarked against experiments
using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
as discussed more detail in the next section.
Scaling down the partial charges reduced the ion binding affinity.
We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
which is only slightly higher than the ECC one ($f_q=0.75$).
Directly applying
the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).

\subsection{Electrometer concept} \label{section:electrometer}
Comparing MD simulation to NMR experiments, we can validate the ion
binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}. From MD simulations the order parameters can be calculated using the relation
\begin{equation}\label{OP}
S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.

The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
\begin{equation}\label{OPchangeEQ}
\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
\end{equation}
where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.

The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).  The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities. This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}. To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}. DHAB\\[0.5cm]
%\vspace{0.5cm}
\chemfig{
  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
}  
\vspace{0.5cm} \\
is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end. Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.

\subsection{Salt concentrations and binding affinities}
When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
while others use atomic absorption spectroscopy and report the salt concentration in
the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
we stick to the latter definition by estimating the salt concentration in the aqueous
bulk region using the farthest point from the lipid bilayer in the aqueous phase.
Note that the former definition was used by some of us previously.~\cite{catte16}.
Although these two definitions provide somewhat different values when applied
to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.

To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}. Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region. In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.  Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
 \begin{equation}\label{surfexcess}
  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
\end{equation}
Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase,
and $A_b$ is the area of the unit cell in the membrane plane.
The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
because the bilayer has an interface at each of the two leaflets.

\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors. NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq. \ref{OP}. Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule). Form factors were calculated using a relation

@@

\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}

@@

First, we present results for bilayers in pure water.
The ECC-POPC and Lipid14 models both reproduce the experimental X-ray scattering form factors
of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
is $\approx$1\AA{} smaller than the experimental one.
The values of the area per lipid of the ECC-POPC model vary slightly
when simulated with different water models (i.e., within the interval of 62.2--66.8 \AA{}, see Table~S2 in SI),
while still being close to the experimentally reported values.
We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
lipid bilayer with a comparable accuracy to other state-of-the-art lipid models~\cite{ollila16}.

>>>>

\begin{abstract}
  % insert abstract here
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. Current estimates of binding affinities and
stoichiometries of cations are, however, controversial due to limitations in the available experimental and
theoretical methods. Experimentally one can assess these parameters by titrating
several membrane properties as a function of the ion concentration. However, the
interpretation of the experiments relies on theoretical models, which give
the best conception of the ion binding as direct experiments are not available. 
Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
resolution, therefore offering all the necessary information to interpret
experimental data without the need to resort to simplified models. However, the accuracy
of the available lipid models when interacting with ions is not sufficient for
such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
polarization as a mean field correction, known as the electronic continuum correction (ECC).
This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid headgroup to a bound positive charge, and the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens,
in agreement with the previous interpretations of the experimental spectroscopic data.
The new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities, suggesting more complex binding stoichiometry than simple models
previously used to interpret the NMR data.
The results of this work pave the way to quantitative MD simulations of complex
biochemical systems with realistic electrostatic interactions in the vicinity of cellular
membranes.
\end{abstract}

@@

\section{Introduction}
Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}. Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}. Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface. For example, Ca$^{2+}$  is crucial in neural signal propagation. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013Ca2}. A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive. \todo{Hector: I left one unknown citation for Pavel.}

Direct measurements of ion-membrane interactions in real biological systems are difficult. 
Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}. The details of ion binding are, however, not fully consistent in the literature. Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mainly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers. On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}. On the other hand, most atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}. For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
%{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}

Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies. A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept~\cite{seelig87}. Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results. 
Analyzing massive amounts of data collected by an open collaboration method, 
it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.

In this work, we improve the cation binding in a POPC bilayer in MD simulations 
via the implicit inclusion of the electronic polarizability in the polar region of phospholipids. 
For this, we apply an electronic continuum correction (ECC)~\cite{leontyev11}, 
in which scaling of the atomic partial charges is used to account for the missing electronic polarizability in MD simulations. 
Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}. As a starting template for our POPC model, we will use the Lipid14~\cite{dickson14}. This force field provides one of the best available descriptions of cation binding \cite{catte16}. 
The newly developed ECC-POPC model reproduces the experimentally
measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while significantly improving the membrane binding affinities to sodium and calcium cations.

@@

\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
The lack of electronic polarizability in standard MD simulation force fields has been considered a serious issue since the early days of lipid bilayer simulations. In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by including the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}. Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}. However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.

According to ECC, the electronic polarizability can be modeled in non-polarizable MD in a mean-field way by 
embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}. Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
\begin{equation}
Q^{ECC} = f_q \cdot Q
\end{equation}
for the ECC corrected charges in simulation. 
Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$. This scaling factor successfully improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
It is important to note that the 
high frequency dielectric constant 
is around 2 for almost any biologically relevant environment \cite{leontyev11}.
The dielectric discontinuity in a lipid bilayer arises
from the orientational polarization of the molecules. 
Thus, the same correction for the electronic polarizability can be 
applied throughout the lipid bilayer interface.

While using a scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case. Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables. Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}. Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}. Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$. Instead, a consistent scaling factor would lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).

Here we develop ECC-POPC lipid model that accurately describes the binding 
of sodium and calcium ions to the lipid bilayer. 
The Lipid14~\cite{dickson14} force field parameters 
(available in a Gromacs format from Ref. \citenum{lipid14files}) were used as a starting 
point. 
This model provides the best response of the head group to ions among the available 
lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
has a relatively realistic head group, glycerol backbone, and acyl chain structure~\cite{dickson14,botan15}.
We applied the ECC correction 
to the Lipid14 model of POPC by scaling of 
partial charges of the head group, glycerol 
backbone, and carbonyl regions. 
These are the polar parts of phospholipids which can 
contribute to the cation binding.

To reproduce the experimental ion binding affinities,
we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
The ion binding affinity was evaluated against experiments
using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
as discussed more detail in the next section.
Scaling down the partial charges reduced the ion binding affinity.
We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
which is only slightly higher than the theoretical one ($f_q=0.75$).
Directly applying
the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
reduced hydration of the lipid head group region after scaling charges which effectively
reduces head group polarity. We solve this difficulty by reducing the effective radii of
the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a
factor of $f_\sigma = 0.89$. The same solution was taken for the ECC-ions in aqueous
solvent \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the $\sigma$ parameters,
the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).

\subsection{Electrometer concept} \label{section:electrometer}
Comparing MD simulation to NMR experiments, we can validate the ion
binding affinity in lipid bilayer simulations using the ''electrometer concept''~ \cite{seelig87, catte16}.
This method relies on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
%Comparing this response is an excellent way to validate force fields as both NMR and MD can accurately determine order parameters ~\cite{catte16, ollila16}.
The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}. From MD simulations the order parameters can be calculated using the definition
\begin{equation}\label{OP}
S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C-H bond and membrane normal. Angular brackets point to the average over all sampled configurations.

The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
\begin{equation}\label{OPchangeEQ}
\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
\end{equation}
where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and position of the bound charge.

The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).  The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities. This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}. To calibrate the response of the head group order parameter to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}. DHAB\\[0.5cm]
%\vspace{0.5cm}
\chemfig{
  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
}  
\vspace{0.5cm} \\
is a cationic surfactant having two acyl chains and bearing a unit charge in the hydrophilic end. Thus, it is expected to locate in the bilayer similarly to the phospholipids, and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.

\subsection{Salt concentrations and binding affinity}
NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
while others use atomic absorption spectroscopy and report the salt concentration in
the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
%we use the latter definition. In particular,
we measure the latter concentration in the aqueous
bulk region using the farthest point from the lipid bilayer in the aqueous phase.
Notice that the former concentration was used by Catte et al.~\cite{catte16}.
Although these two definitions vary when applied
to Ca$^{2+}$, they do not significantly affect the conclusions of this work.

To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}. Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing plane between the membrane interior and the water bulk region. In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.  Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
 \begin{equation}\label{surfexcess}
  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
\end{equation}
Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase,
and $A_b$ is the area of the unit cell in the membrane plane.
The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
because the bilayer has an interface at each of the two leaflets.

\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and \mbox{x-ray} scattering experiments by the scattering form factors. NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq. \ref{OP}. Scattering form factors validate the dimensions of the lipid bilayer (e.g., thickness and area per molecule). Form factors were calculated using

@@

\subsection{Structural parameters of pure ECC-POPC model membrane are in agreement with experiments}

@@

First, we present results for bilayers in pure water.
The ECC-POPC and Lipid14 models both reproduce the experimental x-ray scattering form factors
of a POPC bilayer with a comparable accuracy in Fig.~\ref{simVSexpNOions}.
The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
is $\approx$1\AA{} smaller
The values of the area per lipid of the ECC-POPC model vary
when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI),
while still being close to the experimentally reported values.
We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
lipid bilayer with the comparable accuracy to other state of the art lipid models~\cite{ollila16}.
